---
title: "HW 2 Student"
author: "Anna Fetter"
date: "9/27/2023"
output: 
  html_document:
    number_sections: true
---

This homework is meant to illustrate the methods of classification algorithms as well as their potential pitfalls.  In class, we demonstrated K-Nearest-Neighbors using the `iris` dataset.  Today I will give you a different subset of this same data, and you will train a KNN classifier.  

```{r, echo = FALSE}
set.seed(123)
library(class)

df <- data(iris) 

normal <-function(x) {
  (x -min(x))/(max(x)-min(x))   
}

iris_norm <- as.data.frame(lapply(iris[,c(1,2,3,4)], normal))

subset <- c(1:45, 58, 60:70, 82, 94, 110:150)
iris_train <- iris_norm[subset,] 
iris_test <- iris_norm[-subset,] 

iris_target_category <- iris[subset,5]
iris_test_category <- iris[-subset,5]


```

#
Above, I have given you a training-testing partition.  Train the KNN with $K = 5$ on the training data and use this to classify the 50 test observations.  Once you have classified the test observations, create a contingency table -- like we did in class -- to evaluate which observations your algorithm is misclassifying.   

```{r}
set.seed(123)
#STUDENT INPUT
pr <- knn(iris_train, iris_test, cl = iris_target_category, k = 5)
tab <-table(pr, iris_test_category)
tab
```

#

Discuss your results.  If you have done this correctly, you should have a classification error rate that is roughly 20% higher than what we observed in class.  Why is this the case? In particular run a summary of the `iris_test_category` as well as `iris_target_category` and discuss how this plays a role in your answer.  

*STUDENT INPUT* 

The error rate is higher because we didn't randomly partition the data. The subset was explicitly called out and included blocks of data, so the testing set was not representative of the training data. The summary data reveals that the spread of the testing data is narrower than the spread of the training data. This can be fixed by randomly partitioning the training and testing data.
```{r}
summary(iris_train)
summary(iris_test)
```

Choice of $K$ can also influence this classifier.  Why would choosing $K = 6$ not be advisable for this data? 

*STUDENT INPUT*

#K=6 is a bad classifier because 6 is divisible by the number of categories, which is 3 in this case. K needs to be indivisible by the number of categories.

Build a github repository to store your homework assignments.  Share the link in this file.  

*STUDENT INPUT*
https://github.com/aefetter/stor390hw2
